---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  warning = FALSE,
  message = FALSE
)
```

# partialling.out

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/partialling.out)](https://CRAN.R-project.org/package=partialling.out)
<!-- badges: end -->

Partialling out is a package that allows to generate residualised variables of already existing linear or fixed effects models. So far it works with `lm`, `felm` (`lfe`package) and `feols`  (`fixest` package) following the Frisch-Waugh-Lovell theorem, as explained in Lovell ([2008](doi:10.3200/JECE.39.1.88-91)). It follows a logic similar to the `fwlplot` package with three main differences.
- It uses an already existing model instead of a formula.
- Works with `lm` and `felm` objects alongside `feols`.
- Returns a data.frame, thus offering more freedom of what to do with the results.


## Installation

You can install the development version of partialling.out from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("marcboschmatas/partialling.out")
```



## Examples

Let's make a simple linear model first.


```{r}
library(partialling.out)
library(tinytable)
library(tinyplot)
library(palmerpenguins)

model <- lm(bill_length_mm ~  bill_depth_mm + species, data = penguins)
summary(model)
```

Using the `partialling_out` function, you can get the residualised variable of interest (bill length) and of the first explanatory variable (bill_length), i.e. it would return the residuals of the following two regressions.

```{r eval = FALSE}

modely <- lm(bill_length_mm ~ species, data = penguins)

modelx <- lm(bill_depth_mm ~ species, data = penguins)

```



```{r}
res <- partialling_out(model, data = penguins)

tt(head(res))
```

Accordingly, the coefficient of `res_bill_depth_mm` in the model `lm(res_bill_length_mm ~ res_bill_depth_mm)` will be the same of the coefficient of `bill_depth_mm` in the original model.

```{r}
resmodel <- lm(res_bill_length_mm ~ res_bill_depth_mm, data = res)

print(c(model$coefficients[2], resmodel$coefficients[2]))
```


If `both` is set to `FALSE`, the function will return the actual Y values and the residualised X values.

```{r}
tt(head(partialling_out(model, penguins, both = FALSE)))
```


## Plotting the results

Results can then be displayed in a scatterplot either regular or binned.

```{r fig.height = 8, fig.width = 6}

tinytheme("tufte")
par(mfrow = c(2, 1))
tinyplot(res$res_bill_depth_mm,
         res$res_bill_length_mm); abline(resmodel)

res$qnt <- cut(res$res_bill_depth_mm,
               breaks = quantile(res$res_bill_depth_mm,
                                 probs = seq(0, 1, .02)),
               include.lowest = TRUE)


res_qnt <- aggregate(cbind(res_bill_depth_mm, res_bill_length_mm) ~ qnt,
                     data = res, FUN = mean)

tinyplot(res_qnt$res_bill_depth_mm,
         res_qnt$res_bill_length_mm); abline(resmodel)
```


## Fixed effects models
As stated, the model will also work with `feols` or `felm` models
```{r}
library(fixest)

model_fixest <- feols(bill_length_mm ~ bill_depth_mm | species,
                      data = penguins)


tt(head(partialling_out(model_fixest, data = penguins)))
```


```{r}
library(lfe)

model_lfe <- felm(bill_length_mm ~ bill_depth_mm | species,
                  data = penguins)


tt(head(partialling_out(model_lfe, data = penguins)))
```


## Adding other parameters to the model

Any parameters that could be passed to `lm()`, `feols()`, or `felm()`, can be passed to  `partialling_out()`.

```{r}
model_fixest <- feols(bill_length_mm ~ bill_depth_mm | species,
                      data = penguins, cluster = ~species)


tt(head(partialling_out(model_fixest, data = penguins, cluster = ~species)))
```



## Acknowledgements
To the authors of the [fwlplot](https://github.com/kylebutts/fwlplot) package, Kyle Butts and Grant McDermott, which has provided inspiration and ideas for this project.
To my colleague Andreu Arenas-Jal for his insight and guiding.
