---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  warning = FALSE,
  message = FALSE
)
```

# partialling.out

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/partialling.out)](https://CRAN.R-project.org/package=partialling.out)
[![Codecov test coverage](https://codecov.io/gh/marcboschmatas/partialling.out/graph/badge.svg)](https://app.codecov.io/gh/marcboschmatas/partialling.out)
<!-- badges: end -->

Partialling out is a package that allows to generate residualised variables of already existing linear or fixed effects models. So far it works with `lm`, `felm` (`lfe`package) and `feols`  (`fixest` package) following the Frisch-Waugh-Lovell theorem, as explained in Lovell ([2008](doi:10.3200/JECE.39.1.88-91)). It follows a logic similar to the `fwlplot` package with three main differences.

- It uses an already existing model instead of a formula.

- Works with `lm` and `felm` objects alongside `feols`.

- Returns a data.frame, thus offering more freedom of what to do with the results.


## Installation

You can install the development version of partialling.out from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("marcboschmatas/partialling.out")
```



## Examples

The workflow for `partialling.out` is rather simple: first, create a linear or fixed effects model.

```{r}
library(partialling.out)
library(tinytable)
library(tinyplot)
library(palmerpenguins)

model <- lm(bill_length_mm ~  bill_depth_mm + species, data = penguins)
summary(model)
```

Using the `partialling_out` function, you can get the residualised variable of interest (bill length) and of the first explanatory variable (bill_length), i.e. it would return the residuals of the following two regressions.

```{r eval = FALSE}

modely <- lm(bill_length_mm ~ species, data = penguins)

modelx <- lm(bill_depth_mm ~ species, data = penguins)

```



```{r}
res <- partialling_out(model, data = penguins)

tt(head(res))
```

Accordingly, the coefficient of `res_bill_depth_mm` in the model `lm(res_bill_length_mm ~ res_bill_depth_mm)` will be the same of the coefficient of `bill_depth_mm` in the original model.

```{r}
resmodel <- lm(res_bill_length_mm ~ res_bill_depth_mm, data = res)

print(c(model$coefficients[2], resmodel$coefficients[2]))
```



## Acknowledgements
To the authors of the [fwlplot](https://github.com/kylebutts/fwlplot) package, Kyle Butts and Grant McDermott, which has provided inspiration and ideas for this project.

To my colleague Andreu Arenas-Jal for his insight and guiding.
