---
title: "partialling_out"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{partialling_out}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```



## Getting started

Let's make a simple linear model first.


```{r}
library(partialling.out)
library(tinytable)
library(tinyplot)
library(palmerpenguins)
library(fwlplot)
model <- lm(bill_length_mm ~  bill_depth_mm + species, data = penguins)
summary(model)
```

Using the `partialling_out` function, you can get the residualised variable of interest (`bill_length_mm`) and of the first explanatory variable (`bill_depth_mm`), i.e. it would return the residuals of the following two regressions.

```{r eval = FALSE}

modely <- lm(bill_length_mm ~ species, data = penguins)

modelx <- lm(bill_depth_mm ~ species, data = penguins)

```



```{r}
res <- partialling_out(model, data = penguins)

tt(head(res)) |>
  format_tt(digits = 2) |>
  style_tt(align = "c")
```

Accordingly, the coefficient of `res_bill_depth_mm` in the model `lm(res_bill_length_mm ~ res_bill_depth_mm)` will be the same of the coefficient of `bill_depth_mm` in the original model.

```{r}
resmodel <- lm(res_bill_length_mm ~ res_bill_depth_mm, data = res)

print(c(model$coefficients[2], resmodel$coefficients[2]))
```


If `both` is set to `FALSE`, the function will return the actual Y values and the residualised X values.

```{r}
tt(head(partialling_out(model, penguins, both = FALSE))) |>
  format_tt(digits = 2) |>
  style_tt(align = "c")
```


## Fixed effects models
As stated, the model will also work with `feols` or `felm` models
```{r}
library(fixest)

model_fixest <- feols(bill_length_mm ~ bill_depth_mm | species,
                      data = penguins)

res_fixest <- partialling_out(model_fixest, data = penguins)
tt(head(res_fixest)) |>
  format_tt(digits = 2) |>
  style_tt(align = "c")
```


```{r}
library(lfe)

model_lfe <- felm(bill_length_mm ~ bill_depth_mm | species,
                  data = penguins)

res_lfe <- partialling_out(model_lfe, data = penguins)


tt(head(res_lfe)) |>
  format_tt(digits = 2) |>
  style_tt(align = "c")
```



## Plotting the results

Results can then be displayed in a scatterplot either regular or binned.

```{r fig.height = 8, fig.width = 6}

tinytheme("tufte")
par(mfrow = c(2, 1))


tinyplot(res_fixest$res_bill_length_mm ~ res_fixest$res_bill_depth_mm)
tinyplot_add(res_fixest$res_bill_length_mm ~ res_fixest$res_bill_depth_mm,
             type = "lm")

res_fixest$qnt <- cut(res_fixest$res_bill_depth_mm,
                      breaks = quantile(res_fixest$res_bill_depth_mm,
                                        probs = seq(0, 1, .02)),
                      include.lowest = TRUE)


res_qnt <- aggregate(cbind(res_bill_depth_mm, res_bill_length_mm) ~ qnt,
                     data = res_fixest, FUN = mean)

tinyplot(res_qnt$res_bill_length_mm ~ res_qnt$res_bill_depth_mm)
tinyplot_add(res_fixest$res_bill_length_mm ~ res_fixest$res_bill_depth_mm,
             type = "lm")
```

Plotting all residuals is equivalent to the following approach in `fwlplot`

```{r fig.height = 4, fig.width = 6}

fwlplot(bill_length_mm ~ bill_depth_mm | species, data = penguins)

```


## Adding other parameters to the model

Any parameters that could be passed to `lm()`, `feols()`, or `felm()`, can be passed to  `partialling_out()`.

```{r}
model_fixest <- feols(bill_length_mm ~ bill_depth_mm | species + island,
                      data = penguins, cluster = ~species)


tt(head(partialling_out(model_fixest, data = penguins, cluster = ~species))) |>
  format_tt(digits = 2) |>
  style_tt(align = "c")
```



## Acknowledgements
To the authors of the [fwlplot](https://github.com/kylebutts/fwlplot) package, Kyle Butts and Grant McDermott, which has provided inspiration and ideas for this project.
To my colleague Andreu Arenas-Jal for his insight and guiding.
